Library Parser

### 2024-06-19 studies/library_parser.qmd
undo previous commit

**TODO**: \
- Separate scripts and functions into separate .qmd files/directories. \
- Update `model_rate_emotion_values` nomenclature process for classification papers.\
- Replace rbind with custom bind function throughout

@@

Now we can convert our list into a data frame with the target bibtex fields. For the last field `MODEL_VALIDATION`, we will apply a different regex pattern which matches all characters following the field name `(?<=MODEL_VALIDATION).*` .

```{r}
# create new column containing information between two adjacent target fields for all entries in list
meta_df <- lapply(meta_list, function(x) str_match(paste0(x, collapse = ' '), field_names))
meta_df <- lapply(meta_list, function(x) str_match(paste0(x, collapse = ' '), field_names))

@@

```{r}
## remove bibtext field formatting
# remove curly braces
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '\\{')) 
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '\\},'))
# remove first '=' (from bibtex field )
meta_df <- apply(meta_df, 2, function(x) str_remove(x, '='))
# remove double-commas
#meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, ',,'))
# remove comments
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '%%.*'))
#meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, ' , '))
# remove extra characters in final column
meta_df[, ncol(meta_df)] = str_remove_all(meta_df[, ncol(meta_df)], '\\}')
meta_df <- as.data.frame(meta_df)

@@

included_studies |> dplyr::tibble()
```

@@

```{r}
check_missing <- function(n_empty_fields = 1) 
{
  included_studies[which(
    rowSums(
      sapply(
        included_studies[,1:ncol(included_studies) - 1],
        function(x) grepl("^\\s*$", x))) > n_empty_fields - 1),]
}
```

@@

```{r eval = F, echo = F}
## function to parse output of multiple models nested within a single cell
parse_model_output = function(df)
{ # loop across rows
  # apply function to replace NAs with empty string
  sapply(1:nrow(df), function(x) {
    df$model_rate_emotion_values <- tidyr::replace_na(df$model_rate_emotion_values, '') # detect cells with rbind in them
    if(str_count(df$model_rate_emotion_values[x], "rbind") > 1) {
      # match first model (name between first underscore and rbind)
      model_1_name <- str_match(df$model_rate_emotion_values[x], 
                                "(?<=_).*?(?=rbind)")
      # match subsequent models (name between MODEL_RATE_EMOTION_VALUES_ and rbind)
      model_names <- str_extract_all(df$model_rate_emotion_values[x], 
                                     "(?<=MODEL_RATE_EMOTION_VALUES_).*?(?=rbind)")
      # combine names
      model_names <- unlist(c(model_1_name, model_names))
      # remove special characters
      model_names <- gsub("[^[:alnum:]]", "", model_names)
      # get corresponding values for each model
      model_results <- str_split(df$model_rate_emotion_values[x], 
                "MODEL_RATE_EMOTION_VALUES")
      # get rid of everything up to "rbind"
      model_results <- lapply(model_results, function(x) str_remove(x, 
                                                          '.+?(?=rbind)'))
      # now unlist the models and give them names to add later
      model_results <- unlist(model_results)
       model_results <- paste(model_names, '=', model_results)
      names(model_results) <- model_names
     # return(model_results)
    }
    else { # otherwise, if dealing with non-listed, just return the existing value
      model_results <- df$model_rate_emotion_values[x]
    }
    return(model_results)
  }
    )
  # see whether rbind appears multiple times

@@

```{r}
# low-level function to extract relevant values from named array
model_result_2_df <- function(x) {
  # evaluate the expression
  x <- rlang::eval_tidy(rlang::parse_expr(x))
  # get name of fitted models from column names
  model_names <- colnames(x)
  # print(model_names)
  # split across model name for summary statistic (e.g., mean, sd, etc.)
  model_statistic <- str_split(model_names, '\\.')
  model_statistic <- unlist(lapply(model_statistic, function(x) x[2]))
  # get values
  model_values <- as.numeric(x)
  # names(model_values) <- 'score'
  print(model_values)
  # get additional attributes (feature.data.exp)
  model_attributes <- rownames(x)
  data.frame(model_attributes)
  # get additional model details
  model_attributes <- data.frame(do.call('rbind', str_split(model_attributes, '\\.')))
  names(model_attributes) <- c('library_id', 'model_id', 'feature_id', 'data_id', 'experiment_id')
  print(names(model_attributes))
  # split across '_' to 
  model_measures <- data.frame(do.call('rbind', stringr::str_split(colnames(x), '_')))
  print(model_measures)
  names(model_measures) <- c('dimension', 'measure')
  model_measures_split <- stringr::str_split(model_measures[,'measure'], '\\.')
  model_measures[,'measure'] <- unlist(lapply(model_measures_split, 
                                              function(x) x[1]))
  return(data.frame(
             model_attributes, 
             model_measures, 
             values = model_values,
             statistic = model_statistic))
}

>>>>

**TODO**: 
- Separate scripts and functions into separate .qmd files/directories. \
- Update `model_rate_emotion_values` nomenclature process for classification papers.\
- Replace rbind with custom bind function throughout

@@

Now we can convert our list into a data frame with the target bibtex fields. For the last field `FINAL_NOTES`, we will apply a different regex pattern which matches all characters following the field name `(?<=FINAL_NOTES).*` .

```{r}
Sys.sleep(1)
# create new column containing information between two adjacent target fields for all entries in list
meta_df <- lapply(meta_list, function(x) str_match(paste0(x, collapse = ' '), field_names))
meta_df <- lapply(meta_list, function(x) str_match(paste0(x, collapse = ' '), field_names))

@@

```{r}
Sys.sleep(1)
## remove bibtext field formatting
# remove curly braces
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '\\{')) 
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '\\},'))
# remove first '=' (from bibtex field )
meta_df <- apply(meta_df, 2, function(x) str_remove(x, '='))
# remove double-commas
#meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, ',,'))
# remove comments
meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, '%%.*'))
#meta_df <- apply(meta_df, 2, function(x) str_remove_all(x, ' , '))
# remove extra characters in final column
meta_df[, ncol(meta_df)] = str_remove_all(meta_df[, ncol(meta_df)], '\\}')
meta_df <- as.data.frame(meta_df)

@@

included_studies |> dplyr::tibble() #|> View()
```
# Track studies to discuss
```{r}
included_studies[str_detect(included_studies$final_notes, 'DISCUSS|Discuss|discuss'),] 
```

## Updates following discussion
```{r}
included_studies[str_detect(included_studies$final_notes, 'DISCUSS|Discuss|discuss'),]

resolution_comments <- c(
  'Report total # of features 8904',
  'Report available statistics (new nomenclature solves issue)',
  'Report as many features are there are in MEM n = 260',
  'Report best model results (Rprop)',
  'Report available statistics (new nomenclature solves issue)',
  'Report on proposed model',
  'Report 50 PCA features'
)

included_studies[str_detect(included_studies$final_notes, 'DISCUSS|Discuss|discuss'),]$final_notes <- resolution_comments

# Track Reformatted Extractions

```{r}
included_studies[which(stringr::str_detect(included_studies$model_rate_emotion_values,
                                     'bind_field')),] -> reencoded

message('Completed re-encodings for ', nrow(reencoded), ' studies')

reencoded

@@

```{r}
check_missing_fields <- function(n_empty_fields = 1) 
{
  included_studies[which(
    rowSums(
      sapply(
        included_studies[,1:ncol(included_studies) - 1],
        function(x) grepl("^\\s*$", x))) > n_empty_fields - 1),]
} 
```

@@

```{r eval = F, echo = F}
## function to parse output of multiple models nested within a single cell
get_nested_studies = function(df)
{ # loop across rows
  # apply function to replace NAs with empty string
  sapply(1:nrow(df), function(x) {
    df$model_rate_emotion_values <- tidyr::replace_na(df$model_rate_emotion_values, '') # detect cells with rbind in them
    if(str_count(df$model_rate_emotion_values[x], "rbind") > 1) {
      # match first model (name between first underscore and rbind)
      model_1_name <- str_match(df$model_rate_emotion_values[x], 
                                "(?<=_).*?(?=rbind)")
      # match subsequent models (name between MODEL_RATE_EMOTION_VALUES_ and rbind)
      model_names <- str_extract_all(df$model_rate_emotion_values[x], 
                                     "(?<=MODEL_RATE_EMOTION_VALUES_).*?(?=rbind)")
      # combine names
      model_names <- unlist(c(model_1_name, model_names))
      # remove special characters
      model_names <- gsub("[^[:alnum:]]", "", model_names)
      # get corresponding values for each model
      model_results <- str_split(df$model_rate_emotion_values[x], 
                "MODEL_RATE_EMOTION_VALUES")
      # get rid of everything up to "rbind"
      model_results <- lapply(model_results, function(x) str_remove(x, 
                                                          '.+?(?=rbind)'))
      # now unlist the models and give them names to add later
      model_results <- unlist(model_results)
       model_results <- paste(model_names, '=', model_results)
      names(model_results) <- model_names
     # return(model_results)
    }
    else { # otherwise, if dealing with non-listed, just return the existing value
      model_results <- df$model_rate_emotion_values[x]
    }
    return(model_results)
  }
    )
  # see whether rbind appears multiple times

@@

```{r}
# low-level function to extract relevant values from named array
parse_model_output <- function(x) {
  # evaluate the expression
  x <- rlang::eval_tidy(rlang::parse_expr(x))
  # get name of fitted models from column names
  model_names <- colnames(x)
  # print(model_names)
  # split across model name for summary statistic (e.g., mean, sd, etc.)
  model_statistic <- str_split(model_names, '\\.')
  model_statistic <- unlist(lapply(model_statistic, function(x) x[2]))
  # get values
  model_values <- as.numeric(x)
  # names(model_values) <- 'score'
  print(model_values)
  # get additional attributes (feature.data.exp)
  model_attributes <- rownames(x)
  data.frame(model_attributes)
  # get additional model details
  model_attributes <- data.frame(do.call('rbind', str_split(model_attributes, '\\.')))
  names(model_attributes) <- c('library_id', 'model_id', 'feature_id', 'data_id', 'experiment_id')
  print(names(model_attributes))
  # split across '_' to 
  model_measures <- data.frame(do.call('rbind', stringr::str_split(colnames(x), '_')))
  print(model_measures)
  names(model_measures) <- c('dimension', 'measure')
  model_measures_split <- stringr::str_split(model_measures[,'measure'], '\\.')
  model_measures[,'measure'] <- unlist(lapply(model_measures_split, 
                                              function(x) x[1]))
  return(data.frame(
             model_attributes, 
             model_measures, 
             values = model_values,
             statistic = model_statistic))
}
